\documentclass[11pt]{article}
\usepackage{fancyhdr}
\pagestyle{fancy}
\newcommand\course{CSC 370}
\newcommand\duedate{November 23, 2019}

\lhead{Oliver Tonnesen\\V00885732}
\chead{\textbf{\Large Assignment Journal}}
\rhead{\course\\\duedate}


\usepackage{amsmath}
\usepackage{tikz}


\begin{document}
\noindent
\textbf{Group Name}: Quantum Database Jumpers
\newline
\textbf{Team Leader}: Braydon Horcoff


\section*{Assignment 1}
\subsection*{Sept 9}
I met up with and formed a group with the other four members of our team.


\subsection*{Sept 6 -- Sept 13}
During this time I was recovering from a major surgical operation and was
unable to contribute anything concrete to the assignment.


\subsection*{Sept 16}
I met with the group to catch up with what had been done while I was away,
we began brainstorming ideas for how to represent the scenario we selected in a
more useful model.


\subsection*{Sept 19}
I translated our work up to this point into a rough ER diagram to later be
finalized and drawn up in software for final submission.


\subsection*{Sept 23}
I met with group one more time to finalize the assignment and prepare for it to
be typed up for submission. I also assisted in making some changes to the ER
diagram.


\subsection*{Oct 5}
I helped the group edit the final draft of our submission.


\section*{Assignment 2}
\subsection*{Oct 24}
I met with the group to finalize our preliminary UML diagram and database schema.


\subsection*{Oct 25}
I formatted the UML diagram for submission, reorganized the document and normalized the database.

\section*{Assignment 3}
\subsection*{Nov 8}
I implemented our schema from assignment 2 in postgres, and wrote the base of
the Python code for our demo web app.


\subsection*{Nov 14}
I extended the postgres database with some stored procedures.


\subsection*{Nov 15}
I added capabilities for password encryption to the database, set up a
virtual environment for Python, and started work on the user interace for the
web app.


\subsection*{Nov 20}
I added some fixes to the database schema and the postgres stored procedures.


\subsection*{Nov 21}
I wrote some more postgres functions and added some functionality to the web
app.


\subsection*{Nov 23}
I added some of my group members' queries to our web app, and finalized the
project for submission.

\section*{Assignment 4}
\subsection*{Nov 21}
Worked on some queries for assignment 4.


\end{document}
