\documentclass{article}
\usepackage{tikz,amsmath}
\title{CSC 320 Assignment 1}
\author{Oliver Tonnesen\\V00885732}
\date{January 29, 2019}
\begin{document}
\maketitle
\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\section{} % Section 1
\subsection{} % Section 1.a
\begin{minipage}{\textwidth}
\include{graphics/1a}
\end{minipage}
\subsection{} % Section 1.b
\begin{minipage}{\textwidth}
\include{graphics/1b}
\end{minipage}
\subsection{} % Section 1.c
\begin{minipage}{\textwidth}
\include{graphics/1c}
\end{minipage}
\subsection{} % Section 1.d
\begin{minipage}{\textwidth}
\include{graphics/1d}
\end{minipage}
\newline
To prove minimality, we provide a distinguishing string $w$ for each pair of
states in the given DFA:
\begin{align*}
	\{\{q_1,q_1\}\}\text{ and }\{\{q_2,q_2\},\{q_2,q_1\}\}: 01\\
	\{\{q_1,q_1\}\}\text{ and }\{\{q_4,q_2\},\{q_4,q_1\}\}: 01\\
	\{\{q_1,q_1\}\}\text{ and }\{\{q_3,q_1\}\}: 001\\
	\{\{q_1,q_1\}\}\text{ and }\{\{q_3,q_2\}\}: 001\\
	\{\{q_2,q_2\},\{q_2,q_1\}\}\text{ and }\{\{q_4,q_2\},\{q_4,q_1\}\}: 1\\
	\{\{q_2,q_2\},\{q_2,q_1\}\}\text{ and }\{\{q_3,q_1\}\}: 01\\
	\{\{q_2,q_2\},\{q_2,q_1\}\}\text{ and }\{\{q_3,q_2\}\}: 01\\
	\{\{q_4,q_2\},\{q_4,q_1\}\}\text{ and }\{\{q_3,q_1\}\}: 1\\
	\{\{q_4,q_2\},\{q_4,q_1\}\}\text{ and }\{\{q_3,q_2\}\}: 1\\
	\{\{q_3,q_2\}\}\text{ and }\{\{q_3,q_1\}\}: 1\\
\end{align*}
We clearly see each pair of states is distinguishable, and so our DFA is minimal.
\section{} % Section 2
We construct an NFA recognizing One-off$(L)$ given the DFA recognizing
$L$, $M$.\\
We first define $M'$ to be the same as $M$, but with non-accepting states in
place of accepting states. Our NFA has a start state with $\varepsilon$-transitions
to $M'$'s start state, and to another state with a transition arrow for 0 and 1
going to the start state of $M$. We now create another accepting state with
transition arrows for 0 and 1 coming from every state in $M'$ that was an
accepting state in $M$.
\newline
\newline
We now have an NFA that accepts any string beginning with a 0 or a 1 followed by
a string in $L$, and also accepts any string beginning with a string in $L$
followed by a 0 or a 1. Thus, our NFA accepts One-off$(L)$.
\section{} % Section 3
\subsection{} % Section 3.a
We shall construct an NFA recognizing $A^R$ given the DFA recognizing $A$.
We first change all accepting states in the DFA to non-accepting states in the
NFA. Then, we reverse all transitions in the DFA, and create a new state with
$\varepsilon$-transtions to all former accepting states, and let it be the start
state. The former start state is now the single accepting state. We have now
constructed an NFA recognizing $A^R$, and so $A^R$ is regular.
\subsection{} % Section 3.b
\subsection*{i.} % Section 3.b.i
We showed above that $A^R$ is regular if $A$ is regular, and we know that the
regular languages are closed under concatenation. Thus, if $A$ and $B$ are
regular, then $A^RB^R=\{w^Ry^R|w\in A,y\in B\}$ is regular.
\subsection*{ii.} % Section 3.b.ii
We construct a DFA recognizing the language as follows:
\newline
\newline
Let $D_A=(Q_A,\Sigma_A,\delta_A,q_{0_A},F_A)$ and $D_B=(Q_B,\Sigma_B,\delta_B,q_{0_B},F_B)$
be DFAs recognizing $A$ and $B$, respectively.
\begin{align*}
	&Q=\{(q_i,q_j,x)|q_i\in A,q_j\in B,x\in\{0,1,2\ldots,2n\}\}\\
	&\Sigma=\Sigma_B\cup\Sigma_A\\
	&\delta((q_i,q_j,x),a)=\left\{
	\begin{array}{@{}ll@{}}
		(\delta_A(q_i,a),q_j,x+1)\text{ if $x\equiv0$ (mod 2)}\\
		(q_i,\delta_B(q_j,a),x+1)\text{ if $x\equiv1$ (mod 2)}
	\end{array}\right.\\
	&q_0=(q_{0_A},q_{0_B},0)\\
	&F=\{(q_i,q_j,2n)|q_i\in F_A,q_j\in F_B\}
\end{align*}
Our DFA follows both $D_A$ and $D_B$ in an alternating fashion and accepts only
when it has found an accepting string of length $n$ from both languages.
\section{} % Section 4
\subsection{} % Section 4.a
Only 2.
\subsection{} % Section 4.b
\begin{minipage}{\textwidth}
\include{graphics/4b1}
\end{minipage}
\begin{minipage}{\textwidth}
\include{graphics/4b2}
\end{minipage}
\begin{minipage}{\textwidth}
\include{graphics/4b3}
\end{minipage}
\begin{minipage}{\textwidth}
\include{graphics/4b4}
\end{minipage}
\begin{minipage}{\textwidth}
\include{graphics/4b5}
\end{minipage}
So the regular expression is $\varepsilon0^*1(0(0\cup2)\cup2)^*\varepsilon$.
\end{document}
